---
title: "Developing Quality Metrics"
output: html_notebook
---

This is a r notebook that uses dynamic time warping to align the profile data. Then we used a LOESS regression
to quantify the profile deviation. 
We chose fill 20 speed 20 as our reference sample. 
```{r,warning=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(dtw)
```

## Develop quality metrics
### Data Preprocessing
```{r,warning=FALSE}
# read in the data
folder <- "measurements/profile/profile2/"
file_list <- list.files(path = folder, pattern = "*.csv")

nsample=length(file_list)/6

profile_lst<-list()
for (i in 1:length(file_list)){
    profile<- read.csv(paste(folder, file_list[i],sep=''))
    name<-substr(file_list[i],start=1,stop=nchar(file_list[1])-5)
    assign(name, profile) 
    profile_lst<-c(profile_lst,profile)
}


#omit the NAs
for (i in seq(1,length(profile_lst),by=2)){
  profile_lst[[i]]<-na.omit(cbind(profile_lst[[i]],profile_lst[[i+1]]))[,1]
  profile_lst[[i+1]]<-na.omit(cbind(profile_lst[[i]],profile_lst[[i+1]]))[,2]
}

profile_height<-c()
profile_fill<-c()
profile_speed<-c()
seq1<-seq(1,(3-1)*6+1,by=6)
seq2<-seq(4*6+1,(nsample-1)*6+1,by=6)
sequence<-c(seq1,seq2)
#sequence<-c(1:3,5:length(file_list))  # get rid of 20_20
for (i in sequence){
  profile_height<-c(profile_height,as.numeric(strsplit(file_list[i],"_")[[1]][2]))
  profile_fill<-c(profile_fill,as.numeric(strsplit(file_list[i],"_")[[1]][3]))
  profile_speed<-c(profile_speed,as.numeric(strsplit(file_list[i],"_")[[1]][4]))
}
length(profile_lst)
```

```{r}
# cleaned some bad data
# filter the missing data in 100_60
profile_lst[[2*12+2]]<-profile_lst[[2*12+2]][which(profile_lst[[2*12+1]]>10000)]
profile_lst[[2*12+1]]<-profile_lst[[2*12+1]][which(profile_lst[[2*12+1]]>10000)]
```

### Profile Alignment and Quantifying Errors
Now let's try a loop implementation too return all the alignments of profile 1
```{r}
fulldf<-list()
errordf<-list()
for (j in seq(1,2*5+1,by=2)){
  ref1_ind<-3*12+j
  xRef<-profile_lst[[ref1_ind]] # load xref (profile 20 20 )
  zRef<-profile_lst[[ref1_ind+1]] #load zref
  
  newdf<-list()
  start_ind<-j
  end_ind<-(nsample-1)*12+j

  for (i in seq(start_ind,end_ind,by=12)[-4]){
  X<-profile_lst[[i]]   # load X
  Z<-profile_lst[[i+1]] # load Z 
  
  
  zShift<-max(zRef)-max(Z)
  Z<-Z+zShift               # align in the z direction 
  
  X<-X[which(Z>0)]
  Z<-Z[which(Z>0)]   # trim data with height under 0
  
  xRef<-xRef[which(zRef>100)]
  zRef<-zRef[which(zRef>100)]   # trim data with height under 0
  
  alignment<-dtw(Z,zRef,k=T) # dtw alignment
 
  # take the highest point of ref and find its matching indices with the compared data
  matchdata<-alignment$index1[which(alignment$index2==as.integer(length(Z)/5))]
  matchref<-alignment$index2[which(alignment$index2==as.integer(length(Z)/5))]
  # align in both x and z direction
  zShift<-zRef[matchref]-Z[matchdata]
  xShift<-xRef[matchref]-X[matchdata]

  X<-X+xShift
  Z<-Z+zShift
  
  # trim the left tail of X
  if(max(min(X),min(xRef))==min(xRef)){      # if ref has a shorter left tail, then trim X
  left_start<-which(X>max(min(X),min(xRef)))
  X<-X[left_start]
  Z<-Z[left_start]
  }

  # trim the right tail of X
  if(min(max(X),max(xRef))==max(xRef)){n
  right_start<-which(X<min(max(X),max(xRef)))
  X<-X[right_start]
  Z<-Z[right_start]
  }
  
  newdf<-c(newdf,data.frame(X))
  newdf<-c(newdf,data.frame(Z))
 }

  # run a loess and store the error
  ref<-loess(zRef~xRef,span=0.01)
  error_lst<-c()
  ind<-(nsample-1)*2-1
  for (i in seq(1,ind,by=2)){
   pred<-predict(ref,newdata = newdf[[i]])
   y<-newdf[[i+1]]
   rmse<-sqrt(sum((pred-y)^2))
   error<-rmse/length(y)
   error_lst<-c(error_lst,error)
  }
  fulldf<-c(fulldf,newdf)
  errordf<-c(errordf,data.frame(error_lst))
}
```

```{r}
# before alignment demonstration
ref1_ind<-3*12+1
data_ind<-5*12+1
xRef<-profile_lst[[ref1_ind]] # load xref (profile 20 20 )
zRef<-profile_lst[[ref1_ind+1]] #load zref
plot(xRef,zRef,col="orange",cex=0.2,xlab="X (um)",ylab="Z (um)",asp=1)
points(profile_lst[[data_ind]],profile_lst[[data_ind+1]],cex=0.2,col="blue",asp=1)
legend("topright", legend=c("reference", "data"),
       col=c("orange", "blue"), lty=1:2, cex=0.8)
```

```{r}
# after alignment
j<-1 # which profile
p<-13# the sixth sample
ndf<-(j-1)/2
ref1_ind<-3*12+j
myPlot<-function(index) {points(y=fulldf[[index+1]],x=fulldf[[index]],cex=0.2,col="blue")}
plot(x=profile_lst[[ref1_ind]],y=profile_lst[[ref1_ind+1]],cex=0.2,col="orange",asp=1,,xlab="X (um)",ylab="Z (um)")
lapply(seq(ndf*(nsample-1)*2+p,ndf*(nsample-1)*2+p,by=2),FUN=myPlot)
legend("topright", legend=c("reference", "data"),
       col=c("orange", "blue"), lty=1:2, cex=0.8)
```



```{r}
# find the average error of each profile
meanlst<-c()
for (i in 1:6){
  mean_error<-mean(errordf[[i]])
  meanlst<-c(meanlst,mean_error)
}

# use the three profiles with min error to create quality metrics
min_index<-head(order(meanlst),3)
profile_resid<-c()
for (i in 1:(nsample-1)){
  resid<-mean(errordf[[min_index[1]]][i],errordf[[min_index[2]]][i],errordf[[min_index[3]]][i])
  profile_resid<-c(profile_resid,resid)
}
```




### Import Roughness 
```{r,warning=FALSE}
# read in the rouoghness data
folder <- "measurements/roughness/roughness3/"
file_list <- list.files(path = folder, pattern = "*.csv")

roughnessdf<-data.frame()
for (i in 1:length(file_list)){
    roughness<- read.csv(paste(folder, file_list[i],sep=''),header=F,skip=7)[-1]
    name<-substr(file_list[i],start=1,stop=nchar(file_list[1])-5)
    assign(name, roughness) 
    roughnessdf<-rbind(roughnessdf,roughness)
}
roughnessdf<-roughnessdf[-c(7,8),] # take out the roughness of 20_20 
names(roughnessdf)<-c("Sa","Sz","Str","Spc","Sdr")

# average out the two area
texture<-c()
for(i in seq(1,(nsample-1)*2-1,by=2)){
  mean<-(roughnessdf$Sdr[i]+roughnessdf$Sdr[i+1])/2
  texture<-c(texture,mean)
}

# read height, fill and speed
height<-c()
fill<-c()
speed<-c()
sequence<-c(1:3,5:length(file_list))  # get rid of 20_20
for (i in sequence){
  height<-c(height,as.numeric(strsplit(file_list[i],"_")[[1]][2]))
  fill<-c(fill,as.numeric(strsplit(file_list[i],"_")[[1]][3]))
  speed<-c(speed,as.numeric(strsplit(file_list[i],"_")[[1]][4]))
}

# create quality df
qualitydf<-cbind.data.frame(texture,profile_resid,height,fill,speed)

# export df
write.csv(qualitydf,"qualitydf.csv",row.names = FALSE)
```

